---
title: "Final_Project"
author: "Annabelle"
date: "11/30/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in packages
```{r}
library(here)
library(readxl)
library(skimr) #super nice summary package 
library(car)
library(MASS)
library(tidyverse)
library(corrplot)
library(effects)
library(ResourceSelection) #for testing glm fit
library(DHARMa) # For testing overdispersion
library(kableExtra)
library(stargazer) #tables for models
library(parallelMap) #For classification tree
library(parallel) # For the cross-validating of the decision tree
```


Load data files
```{r}

ALL <- read_excel(here("Final_Project", "PasturedPoultryFarms.xlsx"), sheet = "All")

# For questions 1 & 2: 
Broilers <- read_excel(here("Final_Project", "PasturedPoultryFarms.xlsx"), sheet = "Broilers")

# For question 3: 
FecalSoil <- read_excel(here("Final_Project", "PasturedPoultryFarms.xlsx"), sheet = "FecalSoil")

# For question 5: 
Compositional <- read_excel(here("Final_Project", "PasturedPoultryFarms.xlsx"), sheet = "Compositional")
```

Check out the data 
```{r}
skim(ALL)
skim(Broilers)
skim(FecalSoil)
skim(Compositional)
```


might want to split first column in the All dataset up so have seperate columns with farm label and flock label.. (nope already been done)

### Question 1

Poultry Science researchers were wondering whether presence/absence
of Camplyobacter was predictive of presence/absence of either Salmonella or Listeria, and, similarly,
whether Salmonella presence/absence could be used to predict Listeria presence/absence. Develop
statistical models/tests for all three of these situations and state your conclusions. 



```{r}

# BacS ~ BacC
BacS_BacC_glm <- glm(Salmonella ~ Campylobact, family = binomial(link = logit), data = Broilers)
summary(BacS_BacC_glm)

plot(Broilers$Campylobact, Broilers$Salmonella, main="Fit of data", xlim = c(0,1))
points(Broilers$Campylobact,pnorm(BacS_BacC_glm$coef[1]+Broilers$Campylobact*BacS_BacC_glm$coef[2]), col = 2, pch = 20, type = "l")
#x1 <- sort(df3$temp)
x1 <- seq(0, 1, by = .0025)
y1 <- pnorm(BacS_BacC_glm$coef[1]+x1*BacS_BacC_glm$coef[2])
points(x1, y1, col = 5, pch = 50, type = "l")
#points(x1, y1, col = 4, pch = 20, type = "p")



# BacL ~ BacC
BacL_BacC_glm <- glm(Listeria ~ Campylobact, family = binomial(link = logit), data = Broilers)
summary(BacL_BacC_glm)

plot(Broilers$Campylobact, Broilers$Listeria, main="Fit of data", xlim = c(0,1))
points(Broilers$Listeria,pnorm(BacL_BacC_glm$coef[1]+Broilers$Campylobact*BacL_BacC_glm$coef[2]), col = 2, pch = 20, type = "l")
#x1 <- sort(df3$temp)
x1 <- seq(0, 1, by = .0025)
y1 <- pnorm(BacL_BacC_glm$coef[1]+x1*BacL_BacC_glm$coef[2])
points(x1, y1, col = 5, pch = 50, type = "l")
#points(x1, y1, col = 4, pch = 20, type = "p")




# BacL ~ BacS
BacL_BacS_glm <- glm(Listeria ~ Salmonella, family = binomial(link = logit), data = Broilers)
summary(BacL_BacS_glm)

plot(Broilers$Salmonella, Broilers$Salmonella, main="Fit of data", xlim = c(0,1))
points(Broilers$Listeria,pnorm(BacL_BacS_glm$coef[1]+Broilers$Campylobact*BacL_BacS_glm$coef[2]), col = 2, pch = 20, type = "l")
#x1 <- sort(df3$temp)
x1 <- seq(0, 1, by = .0025)
y1 <- pnorm(BacL_BacS_glm$coef[1]+x1*BacL_BacS_glm$coef[2])
points(x1, y1, col = 5, pch = 50, type = "l")
#points(x1, y1, col = 4, pch = 20, type = "p")

```

Notes: 
- Only intercepts are significant
- Co-occurrence model as glm 



```{r}
mod_Broilers <- Broilers
new <- mod_Broilers %>% filter(Campylobact == 1)
dim(new) #841
new1 <- new %>% filter(Salmonella == 1)
dim(new1) #151 


df1 <- mod_Broilers %>% pivot_longer(c(Campylobact, Salmonella, Listeria), names_to = 'Bac', values_to = 'Count')

#Given Bac C is present, how many 1s and 0s of BacL and BacS are there 

Campylobact_1 <- mod_Broilers %>% filter(Campylobact == 1) 
colSums(Campylobact_1[,c(8:10)])

Campylobact_0 <- mod_Broilers %>% filter(Campylobact == 0) 
colSums(Campylobact_0[,c(8:10)])
dim(Campylobact_0)#will give correct number of Campylobact = 0 [[1049]]


Campylobact_1 <- mod_Broilers %>% filter(Campylobact == 1) 
colSums(Campylobact_1[,c(8:10)])

Campylobact_0 <- mod_Broilers %>% filter(Campylobact == 0) 
colSums(Campylobact_0[,c(8:10)])
dim(Campylobact_0)#will give correct number of Campylobact = 0 [[1049]]

#Given Bac S is present, how many 1s and 0s of BacL are there 
Salmonella_1 <- mod_Broilers %>% filter(Salmonella == 1) 
colSums(Salmonella_1[,c(8:10)])

Salmonella_0 <- mod_Broilers %>% filter(Salmonella == 0) 
colSums(Salmonella_0[,c(8:10)])
dim(Salmonella_0)#will give correct number of Campylobact = 0 [[1567]]


# Back calculated the rest of the tables which results in.... 

#BacS ~ BacC
BacS_given_BacC <- data.frame (BacC0  = c(877, 690),
                  BacC1 = c(172, 151)
                  )
rownames(BacS_given_BacC) <- c("BacS0", "BacS1")


#BacL ~ BacC
BacL_given_BacC <- data.frame (BacC0  = c(891, 715),
                  BacC1 = c(158, 126)
                  )
rownames(BacL_given_BacC) <- c("BacL0", "BacL1")


#BacL ~ BacS
BacL_given_BacS <- data.frame (BacS0  = c(1326, 280),
                  BacS1 = c(241, 43)
                  )
rownames(BacL_given_BacS) <- c("BacL0", "BacL1")


chisq.test(BacS_given_BacC)
chisq.test(BacL_given_BacC)
chisq.test(BacL_given_BacS)

matBacS_given_BacC <- as.matrix(BacS_given_BacC)
corrplot(matBacS_given_BacC, is.cor = FALSE)
chsq <- chisq.test(BacS_given_BacC)
chsq$residuals
corrplot(chsq$residuals, is.cor = FALSE)
contrib <- 100*chsq$residuals^2/chsq$statistic
round(contrib, 3)
corrplot(contrib, is.cor = FALSE)
# printing the p-value
chsq$p.value
# printing the mean
chsq$estimate


## Try to make bar graph representing raw counts in chi square counts 
df2 <- df1[-which(is.na(df1$Count )),] # remove NAs

df2 %>% ggplot((aes(factor(Bac), fill = factor(Count)))) + 
  geom_bar(position = position_dodge(preserve = "single")) + 
  ggtitle("Count of the presence and absence of pathogens") + 
  scale_y_continuous(name = "Count")+
  labs(fill = "Present or Absent") + 
  scale_x_discrete(name = "Name of Pathogen")

### FIGRUE 1 

ggsave(filename = "Figure1_Q1_CountPathogens.png" , width = 10, heigh = 5, device = 'png')


```





Need to make Tables for raw counts for chi square analysis 
```{r}

# Back calculated the rest of the tables which results in.... 

#BacS ~ BacC
BacS_given_BacC <- data.frame (BacC_0  = c(877, 690, 1567),
                  BacC_1 = c(172, 151, 323),
                  Total = c(1049, 841, 1890)
                  )
rownames(BacS_given_BacC) <- c("BacS_0", "BacS_1", "Total")


#BacL ~ BacC
BacL_given_BacC <- data.frame (BacC0  = c(891, 715, 1606),
                  BacC1 = c(158, 126, 284), 
                  Total = c(1049, 841, 1890)
                  )
rownames(BacL_given_BacC) <- c("BacL0", "BacL1", "Total")


#BacL ~ BacS
BacL_given_BacS <- data.frame (BacS0  = c(1326, 280, 1606),
                  BacS1 = c(241, 43, 284),
                  Total = c(1567, 323, 1890)
                  )
rownames(BacL_given_BacS) <- c("BacL0", "BacL1", "Total")

```




### Question 2

Exploratory analysis - correlation matrix 
Note - this will mess up knit because functions are defined below... 
```{r}


#Fit need to create dummy variables for response variables.. 

Broilers2 <- Broilers

#First remove NAs
Broilers2 <- Broilers2[-which(is.na(Broilers2$SampleType )),] # remove NAs


## Make set up for box plots... 
Broilers4 <- Broilers2 %>% 
  mutate(Feces = case_when(grepl("Feces", SampleType, ignore.case = TRUE) ~ 1))%>% 
  mutate(Soil = case_when(grepl("Soil", SampleType, ignore.case = TRUE) ~ 1))%>% 
  mutate(Layer = case_when(grepl("Ceca", SampleType, ignore.case = TRUE) ~ 1))%>% 
  mutate(WCR_P = case_when(grepl("WCR-P", SampleType, ignore.case = TRUE) ~ 1))%>% 
  mutate(WCR_F = case_when(grepl("WCR-F", SampleType, ignore.case = TRUE) ~ 1)) %>% 
  replace(is.na(.), 0) 



#.. Not sure if this is worth doing for binary data.. 

#pairs(FecalSoil2[c(7,11:16)], lower.panel=panel.smooth, cex = .8, pch = 21, bg="steelblue",
   #    diag.panel=panel.hist, cex.labels = 1.2, font.labels=2, upper.panel=panel.cor)







df2 %>% ggplot((aes(factor(Bac), fill = factor(Count)))) + 
  geom_bar(position = position_dodge(preserve = "single"))

Broilers4 %>% 
  ggplot(aes(x=factor(SampleType), fill = factor(Campylobact) )) + 
  geom_bar(stat = "count", position = "fill") + 
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Relationship Between Sample Type and Campylobact") + 
  scale_x_discrete(name = "Sample Type") +
  labs(fill = "Present (1) or Absent (0)")


ggsave(filename = "Figure2a_Q2_Type_BacC.png" , width = 10, heigh = 5, device = 'png')


Broilers4 %>% 
  ggplot(aes(x=factor(Month), fill = factor(Campylobact) )) + 
  geom_bar(stat = "count", position = "fill") + 
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Relationship Between Month and Campylobact") + 
  scale_x_discrete(name = "Month") +
  labs(fill = "Present (1) or Absent (0)")

ggsave(filename = "Figure2b_Q2_Month_BacC.png" , width = 10, heigh = 5, device = 'png')


```



```{r}

Month_type_BacC <- Broilers %>% group_by(Month, SampleType) %>% summarise(countCampylobact = sum(Campylobact)) 

Type_Month_BacC <- Broilers %>% group_by(SampleType, Month) %>% summarise(countCampylobact = sum(Campylobact)) 


#find denominator 
denom1 <- Broilers %>% group_by(Month, SampleType) %>% summarize(count=n()) 
denom2 <- Broilers %>% group_by(SampleType, Month) %>% summarize(count=n()) 


#Add to other dataset 
Month_type_BacC <- full_join(Month_type_BacC, denom1)
Type_Month_BacC <- full_join(Type_Month_BacC, denom2)

Month_type_BacC <- Month_type_BacC %>% mutate(CampylobactAbsent = count - countCampylobact) 
Type_Month_BacC <- Type_Month_BacC %>% mutate(CampylobactAbsent = count - countCampylobact) 

Month_type_BacC <- Month_type_BacC[-which(is.na(Month_type_BacC$Month )),] # remove NAs
Type_Month_BacC <- Type_Month_BacC[-which(is.na(Type_Month_BacC$Month )),] # remove NAs

df_prob2 <- Month_type_BacC %>% 
  unite(ID, SampleType, Month, sep = '_')

df_prob2 <- t(df_prob2)
#colnames(df_prob2) <- df_prob2[1,]
df_prob2 <- df_prob2[-c(1),]

Broilers4 <- Broilers2
Broilers4 <- Broilers4 %>% 
  unite(ID, SampleType, Month, sep = '_')

Broilers5 <- Broilers4 %>% full_join(Broilers2)


Broilers5 %>% 
  ggplot(aes(x=factor(ID),y=as.numeric(Campylobact), colour = factor(SampleType))) + 
  geom_boxplot() +
  coord_flip()+
  ggtitle("Relationship between Type-Month and Ecoli") + 
  scale_x_discrete(name = "Unique Type-Month ID") + 
  scale_y_discrete(name = "Campylobact") + 
  labs(color='Sample Type') 


Broilers5 %>% 
  ggplot(aes(x=factor(ID), fill = factor(Campylobact) )) + 
  geom_bar(stat = "count", position = "fill") + 
  scale_y_continuous(labels = scales::percent)+   
  coord_flip()+
  ggtitle("Relationship between Type-Month and Campylobact") + 
  scale_x_discrete(name = "Unique Type-Month ID") + 
  labs(color='Sample Type') 


ggsave(filename = "Figure3_Q2_Type_Month_BacC.png" , width = 10, heigh = 5, device = 'png')



### FIGRUE 2
### Save me to plot 
### !!!!!!!!! NEED TO FIX THIS PLOT SHOULD BE FOR BAC C NOT ECOLI
 
```


Fit model 

BacC ~ Type + month 
```{r}



g1 = glm(formula = cbind(Month_type_BacC$CampylobactAbsent, Month_type_BacC$countCampylobact) ~ (Month) + SampleType, family = binomial(link = logit), data = Month_type_BacC)
summary(g1)

trial <- Month_type_BacC
trial$Month <- factor(trial$Month)
trial$SampleType <- factor(trial$SampleType)

g1 = glm(formula = cbind(CampylobactAbsent, countCampylobact) ~ (Month) + SampleType, family = binomial(link = logit), data = Month_type_BacC)
summary(g1)

#g2 = glm(formula = cbind(Type_Month_BacC$CampylobactAbsent, Type_Month_BacC$countCampylobact) ~ Month + SampleType, family = binomial, data = Type_Month_BacC)
#summary(g2)


## Same results for g3 as g4
g3 = glm(formula = cbind(Month_type_BacC$CampylobactAbsent, Month_type_BacC$countCampylobact) ~ factor(ID), family = binomial(link = logit), data = df_prob2)
summary(g3)

## Same results for g3 as g3
factoredID <- factor(Broilers4$ID)
g4 = glm(formula = Campylobact ~ factoredID, family = binomial(link = logit), data = Broilers4)
summary(g4)

stargazer(g4, type = "html", title = "Joint-Model for effect of Type-Month on the presence or absence of Campylobact", align = TRUE, out = "Q2_model_stargazer.html" )

## Okay so turns out you get the same model regardless of which order you subset in 

Month_type_BacC <- Month_type_BacC %>% mutate(prop = countCampylobact/count)
#x1 <- (Type_Month_BacC$countCampylobact)


### This graph isn't plotting correctly.... 
plot(Month_type_BacC$countCampylobact,Month_type_BacC$prop, main="Fit of data", xlim = c(0,20))
x1 <- seq(0, 20, by = .5)
y1 <- plogis(g1$coef[1]+x1*g1$coef[2])
#plot(x1, Type_Month_BacC$prop, xlim = c(0,90))
points(x1, y1, col = 5, pch = 20, type = "l", lwd=3)




```

check fit of model 

~ 35 interaction terms in the model

```{r}

## VIFs

vif(
  glm(formula = cbind(Type_Month_BacC$CampylobactAbsent, Type_Month_BacC$countCampylobact) ~ Month + SampleType, family = binomial, data = Type_Month_BacC)
)

# Can't test VIFs for this one because have fewer than 2 predictors.. 
#vif(
#glm(formula = Campylobact ~ factoredID, family = binomial(link = logit), data = Broilers4)
#)


plot(allEffects(g1))

plot(allEffects(g4))

#### How well does our model fit the data? A simple check is a χ2 test for the standardized residuals
## sample size
nn <- 1890 ##dim(Broilers4)
## residuals
ee <- residuals(g4, type = "response")
## fitted values
y_hat <- fitted(g4)
## standardized residuals
rr <- ee / (y_hat * (1 - y_hat))
## test stat
x2 <- sum(rr)
## chi^2 test
pchisq(x2, nn - length(coef(g4)) - 1, lower.tail = FALSE) 
## The p-value is large so we detect no lack of fit





plot(g4) # have problems with qq plot and leverage plot.. 


# If you are running a binary logistic model, you can also run the Hosmer Lemeshow Goodness of Fit test on your glm() model. Using the ResourceSelection library.
hoslem.test(g4$y, g4$fitted)  ## The p-value is large so we conclude an adequate fit


### For the final model in (d), draw the index plot of residuals for the three types of residuals we mentioned in the lecture notes.
 # So the standard residuals, pearson residuals, deviance residuals 

standard = g4$y -  g4$fitted #observed - predicted 
pearson <-  residuals(g4, "pearson")
deviance <-  residuals(g4, "deviance")

plot(standard, col = 1, main = "Standard Residuals")
abline(h=0,col=1)

plot(pearson, col = 2, main = "Pearson Residuals")
abline(h=0,col=2)

plot(deviance, col = 3, main = "Deviance Residuals")
abline(h=0,col=3)




# Leaverge: 
x <- model.matrix(g4)
lev <- hat(x)
plot(lev)
abline(h = 2*5/20) #dim(cins)

leveragePlot(g4)

cooks.distance(g4)


### Overdispersion 

g4$deviance/g4$df.residual ## Should be 1

1-pchisq(g4$deviance,g4$df.residual)

dispersiontest(g4,trafo=1)  ## Poisson GLMs can be tested


sim_fmp <- simulateResiduals(g4, refit=T) 
testOverdispersion(sim_fmp)
testDispersion(sim_fmp)



## function to calculate dispersion parameter
c_hat <- function(obs, model) {
  X2 <- sum((obs - fitted(model))^2 / fitted(model))
  return(X2 / (length(obs) - length(coef(model))))
}
## dispersion parameter
(c_hat_ted <- c_hat(Broilers4$Campylobact, g4))





#dim(Broilers) ## 1892
nn <- 1892 ## sample size
## overdispersion parameter
c_hat <- deviance(g2) / (nn - 2) ## nn is sample size 
## re-scaled estimates
faraway::sumary(g2, dispersion = c_hat)
#Now compare that with original model 
summary(g2)


```
qq plot notes - "Notice the points fall along a line in the middle of the graph, but curve off in the extremities. Normal Q-Q plots that exhibit this behavior usually mean your data have more extreme values than would be expected if they truly came from a Normal distribution."
But the deviance residuals don't have to be normally distributed for the model to be valid, so the normality / non-normality of the residuals doesn't necessarily tell you anything


reference - https://stats.stackexchange.com/questions/121490/interpretation-of-plot-glm-model
& - https://statmath.wu.ac.at/courses/heather_turner/glmCourse_002.pdf



### Question 3


Find a parsimonious model of best fit, performing residual analyses and checking for the effects of any influential points.

```{r}
FecalSoila <-  FecalSoil

## Remove the 2 SampleIDs which are NAs.. 
FecalSoila <- FecalSoila[-which(is.na(FecalSoila$SampleID)),]

FecalSoila$SampleType <- as.factor(FecalSoila$SampleType)
FecalSoila$AnimalSource <- as.factor(FecalSoila$AnimalSource)

# let's try dummy coding all the cateogrical variables 
FecalSoil2 <- FecalSoila %>% dplyr::filter(grepl('Broiler', AnimalSource)) 
  
FecalSoil2 <- FecalSoila  %>% 
  mutate(Broiler = case_when(grepl("Broiler", AnimalSource, ignore.case = TRUE) ~ 1))%>% 
  mutate(Cattle = case_when(grepl("Cattle", AnimalSource, ignore.case = TRUE) ~ 1))%>% 
  mutate(Layer = case_when(grepl("Layer", AnimalSource, ignore.case = TRUE) ~ 1))%>% 
  mutate(Swine = case_when(grepl("Swine", AnimalSource, ignore.case = TRUE) ~ 1))%>% 
  mutate(Feces = case_when(grepl("Feces", SampleType, ignore.case = TRUE) ~ 1))%>% 
  mutate(Soil = case_when(grepl("Soil", SampleType, ignore.case = TRUE) ~ 1)) %>% 
  replace(is.na(.), 0) # then replace all 0s 


#Change into proportion 
colSums(FecalSoil2[,c(11:16)]) 
FecalSoil2 <- FecalSoil2 %>%
  mutate(BroilerProp = 1260/1630)%>% 
  mutate(CattleProp = 90/1630)%>% 
  mutate(LayerProp = 180/1630)%>% 
  mutate(SwineProp = 100/1630)%>% 
  mutate(FecesProp = 815/1630)%>% 
  mutate(SoilProp = 815/1630)

#Having issues with the class.. need to change to numeric 
FecalSoil2[,c(7,11:16)] <- sapply(FecalSoil2[ ,c(7,11:16)], as.numeric)



#set up
PredictorsOnlyPixel <- FecalSoil2[,c(7)]
PredictAndResponsePixel <- FecalSoil2[,c(11:16)]
PredictAndResponseGrid <- FecalSoil2[,c(11:16)]
  
# put histograms on the diagonal panel	
panel.hist <- function (x,...)					# define a function that says what we want to plot in the diagonal
{
  usr <- par("usr"); on.exit(par(usr))			# not sure what usr is for?
  par(usr = c(usr[1:2],0,1.5))
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)			# make the hist 
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col="grey", ...)  # defines what the histogram is going to look like
}

# put correlations on the upper panels,
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- cor(x, y,use="complete.obs")				
  txt <- format(c(r, 0.123456789), digits=digits)[1]
  prefix <- "r = "
  rc <- cor.test(x,y,method = c("pearson"))				## calculate pearsons rho for upper grid
  txt <- paste(prefix,txt,sep="")
  text(0.5, 0.5, txt, cex = 1)
}

## plot a correlation matrix plot that uses the functions specified above to say what to plot where
      ## this was taken directly from website and still not plotting r values for all 
#pairs(PredictAndResponsePixel[1:6], lower.panel=panel.smooth, cex = .8, diag.panel=panel.hist, cex.labels = 1.2, font.labels=2, upper.panel=panel.cor)

##Pixel level

pairs(FecalSoil2[c(7,11:16)], lower.panel=panel.smooth, cex = .8, pch = 21, bg="steelblue",
       diag.panel=panel.hist, cex.labels = 1.2, font.labels=2, upper.panel=panel.cor)

pdf("cormat_model_Q3")
pairs(FecalSoil2[c(7,11:16)], lower.panel=panel.smooth, cex = .8, pch = 21, bg="steelblue",
       diag.panel=panel.hist, cex.labels = 1.2, font.labels=2, upper.panel=panel.cor)
while (!is.null(dev.list()))  dev.off()

### try and figure out how to change lower.panel=panel.smooth to box plots (or to add jitter) right now, just using a predefined function 


FecalSoil2 %>% ggplot(aes(x=factor(AnimalSource),y=EcoliLog10)) + 
  geom_boxplot()  + 
  ggtitle("Relationship Between Ecoli Level and Animal Source Type") + 
  scale_x_discrete(name = "Animal Source Type") 


ggsave(filename = "Figure_Q4a_EcoliAnimalSource_Boxplot.png" , width = 10, heigh = 5, device = 'png')


FecalSoil2 %>% ggplot(aes(x=SampleType,y=EcoliLog10)) + 
  geom_boxplot() + 
  ggtitle("Relationship Between Ecoli Level and Sample Type") + 
  scale_x_discrete(name = "Sample Type") 



ggsave(filename = "Figure_Q4b_EcoliSampleType_Boxplot.png" , width = 10, heigh = 5, device = 'png')


## FIGURE 4 

```



```{r}

#remove NAs 
FecalSoil <- FecalSoil[-which(is.na(FecalSoil$EcoliLog10 )),] # remove NAs
#Have issues with class in df - need to change EcoliLog10 to numeric 
FecalSoil[,c(7)] <- sapply(FecalSoil[ ,c(7)], as.numeric)


model <- lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoil)
summary(model)

vif(lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoil))


# Perform residual analysis 

r <- model$residuals
f <- model$fitted.values
plot(f, r)

hist(model$residuals)


# Leaverge: 
x <- model.matrix(model)
lev <- hat(x)
plot(lev)
abline(h = 2*5/20) #dim(cins)


plot(model)

```


Addressing influential points.. 

```{r}



## Original model 
model <- lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoil)
summary(model)

# Influential observations
cook1 <- cooks.distance(model)
plot(cook1)

n <- nrow(FecalSoil)
plot(cook, main = "Cooks Distance for Influential Obs")
abline(h = 4/n, lty = 2, col = "steelblue") # add cutoff line


## points 137 and 385 are problematic.. 

## Let's fit three seperate models - one without 137 [0.02334191], one without 385 [0.02220467 ] and one without both 

top_x_outlier <- 1
influential1 <- as.numeric(names(sort(cook1, decreasing = TRUE)[1:top_x_outlier]))

FecalSoila <- FecalSoil[-influential1, ]
modela <- lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoila)
cook2 <- cooks.distance(modela)
plot(cook2)
identify(1:1623, cook2)


#influential2a <- as.numeric(names(sort(cook2, decreasing = TRUE)[1:top_x_outlier]))
# IDs = AB1-2, EE2-5
#influential2b <- influential2a -1
### removing the second most influential point still isn't working.. 
influential <- as.numeric(names(cook1)[(cook1 = 0.02220467)])
FecalSoilb <- FecalSoil %>% filter(!(SampleID == "EE2-5" | SampleID == "EE2-4" | SampleID == "EE2-3"))
modelb <- lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoilb)
cook3 <- cooks.distance(modelb)
plot(cook3)
identify(1:1623, cook3)





### Remove both ... 
# Alternatively, you can try to remove the top x outliers to have a look
top_x_outlier <- 2
influential <- as.numeric(names(sort(cook1, decreasing = TRUE)[1:top_x_outlier]))

FecalSoil4 <- FecalSoil[-influential, ]
modelc <- lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoil4)
cook4 <- cooks.distance(modelc)
plot(cook4)



### Remove all points over threshold

influential <- as.numeric(names(cook1)[(cook1 > (4/1628))]) ## dim(FecalSoil4)
FecalSoil5 <- FecalSoil[-influential, ]
modeld <- lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoil5)
cook5 <- cooks.distance(modeld)
plot(cook5)


### scrap

### Manually calculating cooks distance 

Hmat = hatvalues(model)
Leverage = Hmat/(1 - Hmat)
mse = (model$residuals)^2/var(model$residuals)
CooksD <- (1/6)*(mse)*Leverage



FecalSoil$cookd <- cooks.distance(model)

halfnorm(cook, nlab = 2, ylab = "Cook’s Distance")


countries <- rownames(prob4)
x <- c(1:length(FecalSoil$SampleID)) ## length(x) 1630
cook_base <- cooks.distance(model) ## length(cook_base) 1624
plot(cook_base, type = "n", main = "Cook's Distance base model", cex.main = 1, 
     xlab = "countries", ylab = "Cook's Distance")
segments(x,0,x, cook_base, lwd = 2)
identify(1:1624, cook_base, FecalSoil$SampleID, cex = .75)

```

Present the findings from cook's distance manipulations
```{r}

stargazer(model, modelb, modelc, type = "html", title = "The Effect of Influential Points", align = TRUE, out = "Q3_influential_points_models_stargazer.html" )

plot(cook1, main = "Cook's Distance for Entire Model", cex.main = 1, 
     xlab = "Samples", ylab = "Cook's Distance")
abline(h = 4/n, lty = 2, col = "steelblue") # add cutoff line

plot(cook2, main = "Cook's Distance: Removed One Influential Point", cex.main = 1, 
     xlab = "Samples", ylab = "Cook's Distance")
abline(h = 4/n, lty = 2, col = "steelblue") # add cutoff line

plot(cook4, main = "Cook's Distance: Removed One Influential Point", cex.main = 1, 
     xlab = "Samples", ylab = "Cook's Distance")
abline(h = 4/n, lty = 2, col = "steelblue") # add cutoff line


modeld <- lm(EcoliLog10 ~ factor(SampleType) + factor(AnimalSource), data = FecalSoil5)
cook5 <- cooks.distance(modeld)
plot(cook5)
abline(h = 4/n, lty = 2, col = "steelblue") # add cutoff line

plot(modelc)
hist(modelc$residuals)


x <- model.matrix(modelc)
lev <- hat(x)
plot(lev, main = "Index Plot of Leverages", cex.main = 1, 
     xlab = "Index", ylab = "Leverages")
abline(h = 2*5/1528) #dim(FecalSoil4)

jack <- rstudent(modelc)
plot(jack,ylab="Jackknife Residuals",main="Jackknife Residuals", xlab = "Index")
jack[abs(jack)==max(abs(jack))]


outlierTest(modelc)
## the Bonferroni p-value for the largest (absolute) residual is not statistically significant. While this test is important for identifying a potentially significant outlying observation, it is not a panacea for checking for patterns in outlying data.


#levene's test for constant variance 

```
lightly tailed
getting correlated errors because blocked?



Need to address influential points
```{r}




##### ALso as seen in the q-q plot- there is an issue with outliers 


# Box-Cox transformation.
# Not going to work because some values are 0.. 
#boxcox(model,plotit=T)
#boxcox(model,lambda=seq(0.0,3.0,by=0.05),plotit=T) ### lamda = ~1.5
#Could change "0" values to .000001 or some very small number because data are not really true zeros - there should still be a tiny amount. 




## revised `halfnorm()` from Faraway
## `nlab` gives the number of points to label in the plot
## `labels` can be a character vector the same length as `x`
##          that can be used to label the points in the plot
## `ylab` can be used to labeling of the y-axis
halfnorm <- function(x, nlab = 1, labels = NULL, ylab = "Sorted data") {
  x <- abs(x)
  labord <- order(x)
  x <- sort(x)
  i <- order(x)
  n <- length(x)
  ui <- qnorm((n + 1:n)/(2 * n + 1))
  if(is.null(labels)) {
    labels <- as.character(1:length(x))
  }
  plot(ui, x[i], pch = 16, las = 1,
       xlab = "Half-normal quantiles", ylab = ylab, 
       ylim = c(0, max(x)), type = "n")
  if(nlab < n) {
    points(ui[1:(n - nlab)], x[i][1:(n - nlab)], pch = 16)
  }
  text(x = ui[(n - nlab + 1):n], y = x[i][(n - nlab + 1):n],
       labels = labels[labord][(n - nlab + 1):n])
}

```
Left sqew in q-q plot? 

have bins of variation instead of normally distrbuted







SCRAP



```{r eval= FALSE}

BroilersCampylobact <- Broilers %>% group_by(Farm) %>% group_by(Flock) %>% summarise(countCampylobact = sum(Campylobact)) 
BroilersSalmonella <- Broilers %>% group_by(Farm) %>% group_by(Flock) %>% summarise(countSalmonella = sum(Salmonella)) 
BroilersListeria <- Broilers %>% group_by(Farm) %>% group_by(Flock) %>% summarise(countListeria = sum(Listeria))

BacCount_Flock <- full_join(BroilersCampylobact, BroilersSalmonella) %>% full_join(BroilersListeria)
BacCount_Flock <- BacCount_Flock[-which(is.na(BacCount_Flock$Farm )),] # remove NAs

#find denominator 
denom <- Broilers %>% group_by(Farm) %>% group_by(Flock) %>% summarize(count=n()) # so 45 per group is denominator (there were 2 nas)



BroilersCampylobact <- Broilers %>% group_by(Farm) %>% summarise(countCampylobact = sum(Campylobact)) 
BroilersSalmonella <- Broilers %>% group_by(Farm) %>% summarise(countSalmonella = sum(Salmonella)) 
BroilersListeria <- Broilers %>% group_by(Farm)  %>% summarise(countListeria = sum(Listeria))

BacCount_Farm <- full_join(BroilersCampylobact, BroilersSalmonella) %>% full_join(BroilersListeria)

#find denominator 
BacCount_Farmdenom <- Broilers %>% group_by(Farm)  %>% summarize(count=n()) # Different numbers per farm.. 
#Add to other dataset 
BacCount_Farm <- full_join(BacCount_Farm, BacCount_Farmdenom)

BacCount_Farm <- BacCount_Farm %>% mutate(CampylobactAbsent = count - countCampylobact) %>% 
  mutate(SalmonellaAbsent = count - countSalmonella) %>% 
  mutate(ListeriaAbsent = count - countListeria)

BacCount_Farm <- BacCount_Farm[-which(is.na(BacCount_Farm$Farm )),] # remove NAs


y = c(BacCount_Farm$countSalmonella, BacCount_Farm$SalmonellaAbsent)

g = glm(formula = cbind(BacCount_Farm$countSalmonella, BacCount_Farm$SalmonellaAbsent) ~ BacCount_Farm$countCampylobact, family = binomial)
summary(g)

```




### Question 5


Trying to get 6 different levels of classification 

```{r}

mydata <- Compositional
colnames(mydata) <- c("ID"        ,   "Farm"      ,   "PastureTime" , "Flock"   ,     "Sampletype" ,  "A"  ,      "B"      ,     "C"       ,     "D"       ,     "E" )

mydata <- mydata %>% 
  unite(TimeSample, Sampletype, PastureTime, sep = '_')


mydata <- mydata %>% 
  dplyr::select(TimeSample, A, B,C,D,E)


# Converting logical variables to factors
mydata <- mutate_if(mydata, is.logical, as.factor)


mydataTask <- makeClassifTask(data = mydata, target = "TimeSample")
tree <- makeLearner("classif.rpart")


# Printing available rpart hyperparameters
getParamSet(tree)

## minbucket = 5
##maxdepth = 3

# Defining the hyperparameter space for tuning
treeParamSpace <- makeParamSet(
  makeIntegerParam("minsplit", lower = 5, upper = 20),
  makeIntegerParam("minbucket", lower = 3, upper = 5),
  makeNumericParam("cp", lower = 0.01, upper = 0.1),
  makeIntegerParam("maxdepth", lower = 2, upper = 3))

## results here end up dropping 2 of the categories..   
treeParamSpace <- makeParamSet(
  makeIntegerParam("minsplit", lower = 3, upper = 4),
  makeIntegerParam("minbucket", lower = 3, upper = 5),
  makeNumericParam("cp", lower = 0.01, upper = 0.1),
  makeIntegerParam("maxdepth", lower = 2, upper = 3))


### ### This is what is written up in the report 
## New approach --> 
## results here end up dropping 2 of the categories..   
treeParamSpace <- makeParamSet(
  makeIntegerParam("minsplit", lower = 33, upper = 38), #The minimum number of observations in a given category for this data is 33 (Soil Start) and the maximum number of observations we have is 38 (Fecal End). If we didn’t loose any observations, both should be set to 40 
  makeIntegerParam("minbucket", lower = 33, upper = 38),
  makeNumericParam("cp", lower = 0.01, upper = 0.1),
  makeIntegerParam("maxdepth", lower = 2, upper = 2)) 

### ### This is what is written up in the report 
## New approach --> trying to just get fecal and soil with lower misclassification rate 
treeParamSpace <- makeParamSet(
  makeIntegerParam("minsplit", lower = 33, upper = 111), 
  makeIntegerParam("minbucket", lower = 33, upper = 111),
  makeNumericParam("cp", lower = 0.01, upper = 0.1),
  makeIntegerParam("maxdepth", lower = 2, upper = 6)) 


# Defining the random search
randSearch <- makeTuneControlRandom(maxit = 200)
cvForTuning <- makeResampleDesc("CV", iters = 5)


# Performing hyperparameter tuning
parallelStartSocket(cpus = detectCores()) ## This function doesn't work anymore? 

tunedTreePars <- tuneParams(tree, task = mydataTask,
                            resampling = cvForTuning,
                            par.set = treeParamSpace,
                            control = randSearch)

parallelStop()

tunedTreePars


# Training the final tuned model
tunedTree <- setHyperPars(tree, par.vals = tunedTreePars$x)
tunedTreeModel <- train(tunedTree, mydataTask)

# Plotting the decision tree
treeModelData <- getLearnerModel(tunedTreeModel)
rpart.plot(treeModelData, roundint = FALSE,
box.palette = "BuBn",
type = 5)


# Exploring the model
printcp(treeModelData, digits = 3)


### Cross-validating our decision tree model

# Cross-validating the model-building process
outer <- makeResampleDesc("CV", iters = 5)
treeWrapper <- makeTuneWrapper("classif.rpart", resampling = cvForTuning,
                               par.set = treeParamSpace,
                               control = randSearch)
parallelStartSocket(cpus = detectCores())
cvWithTuning <- resample(treeWrapper, mydataTask, resampling = outer)
parallelStop()

# Extracting the cross-validation result
cvWithTuning


```


Trying to get 2 different levels of classification.. 

```{r}

mydata <- Compositional

library(mlr)

data(Zoo, package = "mlbench")
zooTib <- as_tibble(Zoo)
zooTib
zooTib <- mutate_if(zooTib, is.logical, as.factor)
zooTask <- makeClassifTask(data = zooTib, target = "type") 
tree <- makeLearner("classif.rpart")
getParamSet(tree)

#Let's define the hyperparameter space we want to search over
treeParamSpace <- makeParamSet(makeIntegerParam("minsplit", lower = 5, upper = 20), makeIntegerParam("minbucket", lower = 3, upper = 10), makeNumericParam("cp", lower = 0.01, upper = 0.1), makeIntegerParam("maxdepth", lower = 3, upper = 10)) 

randSearch <- makeTuneControlRandom(maxit = 200) 
cvForTuning <- makeResampleDesc("CV", iters = 5)


library(parallel) ;
library(parallelMap) ;
parallelStartSocket(cpus = detectCores()) 
tunedTreePars <- tuneParams(tree, task = zooTask, resampling = cvForTuning, par.set = treeParamSpace, control = randSearch) ; parallelStop()

tunedTree <- setHyperPars(tree, par.vals = tunedTreePars$x) 
tunedTreeModel <- train(tunedTree, zooTask)

install.packages("rpart.plot"); library(rpart.plot) 
treeModelData <- getLearnerModel(tunedTreeModel)
rpart.plot(treeModelData, roundint = FALSE, box.palette = "BuBn", type = 5)

```




```{r}
library(vegan)


set.seed(2)
community_matrix=matrix(
   sample(1:100,300,replace=T),nrow=10,
   dimnames=list(paste("community",1:10,sep=""),paste("sp",1:30,sep="")))

example_NMDS=metaMDS(community_matrix, # Our community-by-species matrix
                     k=2) # The number of reduced dimensions


```







